---
title: "Regression Models Course Project"
author: "José Manuel Peñas"
date: "28 de julio de 2018"
output:
  pdf_document:
    df_print: kable
    number_sections: yes
    toc: yes
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2);library(latexpdf);library(dplyr)
```
# Analysis

**MTCARS** is the data set. . A new factor column *transmission* is added with *Automatic* and *manual* values instea}d of 0 and 1 of *am* for an easier manipulation later.

```{r cars, results='hide', echo=FALSE}
newcars <- mutate(mtcars, transmission = as.factor(ifelse(am==0,"Automatic","Manual")))
```
Looking at the boxplot at the appendix, it seems clear that Manual transmission cars do consumes less fuel than automatic ones. The quantity will be meassured by linear regression models.

## Simple Model
First, the analysis creates a simple model that studies the relationship between mpg, the outcome, and transmission the predictor.
```{r simple, echo=FALSE}
simpleModel <- lm (mpg ~ transmission,newcars)
summary(simpleModel)
```
The estimate intercept represents the coefficient of Automatic cars (**17.15**) and the *transmissionManual* the difference between automatic and manual: **7.24**. 
The P-value is lower enough to discard the null hypothesis on both cases, so transmission has a relationship with gasoline consumption. 
As you can see on the plot generated on Appendix, the residuals plots and QQ plot are good and don't show abnormals patterns.
The R-Squared indicates that the prediction of mpg only is influenced by 33-36 % by the transmission of the car.
It is mandatory to check with with other variables in a multivariate model.

## Multivariate Model
First, the variables to be included in the model have to be found.
```{r findingVars, echo=FALSE}
# Used mtcars as we have the new char transmission column
cor(mtcars)[1,]
```
The variables to include are the ones whose absolute value is bigger than **0.75**: *cyl*, *disp*,*hp*, *wt*.
Now, the multivariate model can be generated and compared with Anova test.
```{r multimodel, echo=FALSE}
multiModel <- lm (mpg ~ cyl + disp + hp + wt + transmission - 1,newcars)
summary(multiModel)
anova(simpleModel,multiModel)
```
It has bigger R-Squared **0.98** and if you check the plot of the model, the 4 graphics are correct with no abnomal patterns.
The Anova test shows a very important decrease on the Residual Sum of Squares: **720** to **163**

# Conclussions
## Is an automatic or manual transmission better for MPG
So the boxplot as the simple and multivariates models show that the **manual transmission is better** for MPG.    

## Quantify the MPG difference between automatic and manual transmissions
The multivariate model gives **1.55** the difference between automatic and manual transmissions.

# Appendix Figures
## Boxplot:
```{r boxplot, echo=TRUE}
ggplot(newcars,aes(transmission, mpg, color=transmission)) +
      geom_boxplot(outlier.colour="red", outlier.shape=16,
             outlier.size=2, notch=FALSE)
```

## Simple Model:

```{r simpFig, echo=TRUE}
par(mfrow = c(2,2));plot(simpleModel)
```

## Multivariate Model:

```{r multFig, echo=TRUE}
par(mfrow = c(2,2));plot(multiModel)
```

